---
title: "Assignment 1"
author: "pb649@scarletmail.rutgers.edu"
date: "2/06/2023"
output: html_document
---
  
```{r}
library(readr)
#Sparrow Example
Bumpus_sparrows <- read_csv("~/Downloads/Bumpus_sparrows.csv")
Bumpus_sparrows
sparr1 <- Bumpus_sparrows
sparr1
```

## Drop 1st column as it does not contain numeric values
```{r}
sparr<- sparr1[,-1]
sparr
```

## Matrix for Survived sparrows
```{r}
Survivors=sparr[1:21,]
Survivors
```

## Matrix for Non-Survived Sparrows
```{r}
Remnants=sparr[22:49,]
Remnants
```

## Computing the means of Survivors

```{r}
colMeans(Survivors)
```

## Computing the means of Non-Survivors
```{r}
colMeans(Remnants)
```

## Covariance matrice for survivors
## Inference
* A positive covariance indicates that both variables are more likely to be high or low at the same time.
* Alar_extent and Total_length has the highest covarience, while L_keel_sternum and L_humerous has the least.

```{r}
cov(Survivors[,c("Total_length","Alar_extent","L_beak_head","L_humerous","L_keel_sternum")])
```

## Covariance matrice for Non-survivors
## Inference
* A positive covariance indicates that both variables are more likely to be high or low at the same time.
* Covariance of L_beak_head and L_humerous are contradicting to correlation of L_beak_head and L_humerous. 

```{r}
cov(Remnants[, c("Total_length","Alar_extent","L_beak_head","L_humerous","L_keel_sternum")])
```

## Correlation of Survivors 
## Inference:
* If the correlation coefficient value is 1 then it is positive correlation. The nearer the coefficient value to 1 the stronger the correlation is.
* Correlation between L_humerous and Alar_extent is the strongest as the coefficient value is much closer to 1. 
* Correlation between L_keel_sternum  and Alar_extent is weak one as the coefficient value is further away from  1.

```{r}
cor(Survivors[,c("Total_length","Alar_extent","L_beak_head","L_humerous","L_keel_sternum")])
```

## Correlation of Non-Survivors
## Inference:
* If the correlation coefficient value is 1 then it is positive correlation. The nearer the coefficient value to 1 the stronger the correlation is.
* Correlation between L_humerous and L_beak_head is the strongest as the coefficient value is much closer to 1. 
* Correlation between L_keel_sternum and L_beak_head has the smallest coefficient value.

```{r}
cor(Remnants[, c("Total_length","Alar_extent","L_beak_head","L_humerous","L_keel_sternum")])

```
## Boxplot for comparison 
## Inference
* If the length is more then it will not survive. If the length is less then they survive. i.e if they are above median value then they won't survive.

```{r}
library (ggplot2)
ggplot(sparr1, aes(x=Total_length, fill=Survivorship))+geom_boxplot()+ labs(title='Survivorship comparison')

```




